### 2013-07-15 doc/dissertation/document/chapters/nonlinear_problem/nonlinear_problem.tex
fixed grammar in nonlinear intro
adding more to nonlinear benchmarks section

\chapter{Monte Carlo Synthetic Acceleration Methods for the
  Navier-Stokes \\Equations\ }
\label{ch:nonlinear_problem}
Nonlinear transport problems are a common occurrence in single and
multiple physics problems in nuclear engineering. Systems of partial
differential equations such as those that describe fluid flow or more
general transport processes when discretized by conventional methods
yield discrete sets of stiff equations with nonlinearities present in
the variables. Traditionally, such systems have been solved by
linearizing them in a form where the nonlinearities in the variables
are eliminated and more traditional linear methods can be used for
solutions. Often characterized as segregated methods where physics
operators are split and their action on the system approximated in
steps, such methods lack consistency and accuracy in resolving the
nonlinear component of the solution. In the last 30 years, fully
consistent nonlinear methods based on Newton's method have become more
popular and many advances have been made in the nuclear engineering
field to employ these methods.

@@

\subsection{Newton-Krylov Methods\ }
\label{subsec:newton_krylov_methods}
A form of inexact Newton methods, \textit{Newton-Krylov methods} are
nonlinear iterative methods that leverage a Krylov subspace method as
the linear solver for generating the Newton correction
\citep{kelley_iterative_1995}. As we outlined in
Appendix~\ref{ch:linear_problem}, Krylov methods are robust and enjoy
efficient parallel implementations on modern
architectures. Furthermore, their lack of explicit dependence on the
operator make them easier to implement than other
methods. Additionally, although many iterations can become memory
intensive due to the need to store the Krylov subspace for the
orthogonalization procedure, at each nonlinear iteration this cost is
reset as the Jacobian matrix will change due to its dependence on the
solution vector. This means that for every nonlinear iteration, a
completely new linear system is formed for generating the Newton
correction and we can modify the Krylov solver parameters accordingly
to accommodate this. In most nonlinear problems, the Jacobian operator
is generally non-symmetric and therefore either Krylov methods with
long recurrence relations that can handle non-symmetric systems must
be considered or the Newton correction system must be preconditioned
such that the operator is symmetric and short recurrence relation
methods can be potentially be used.

@@

\subsection{Thermal Convection Cavity Problem}
\label{subsec:natural_convection_cavity}
In 1983 a benchmark solution for the natural convection of air in a
square cavity was published \citep{de_vahl_davis_natural_1983} as
shown in Figure~\ref{fig:natural_convection_cavity} for the solution
of the energy, mass, and momentum equations.
\begin{figure}[t!]
  \begin{center}
    \scalebox{1.5}{
      \input{chapters/nonlinear_problem/natural_convection_cavity.pdftex_t} }
  \end{center}
  \caption{\textbf{Problem setup for the natural convection cavity
      benchmark.} \textit{Dirichlet conditions are set for the
      temperature on the left and right while Neumann conditions are
      set on the top and bottom of the Cartesian grid. The temperature
      gradients will cause buoyancy-driven flow. Zero velocity
      Dirichlet conditions are set on each boundary. No thermal source
      was present.}}
  \label{fig:natural_convection_cavity}
\end{figure}
In this problem, a rectilinear grid is applied to the unit square. No
heat flow is allowed out of the top and bottom of the square with a
zero Neumann condition specified. Buoyancy driven flow is generated by
the temperature gradient from the cold and hot Dirichlet conditions on
the left and right boundaries of the box. By adjusting the Rayleigh
number of the fluid (and therefore adjusting the ratio of convective
to conductive heat transfer), we can adjust the influence of the
nonlinear convection term in Eq~(\ref{eq:ns_momentum}). In Shadid's
work, Rayleigh numbers of up to \sn{1}{6} were used for this benchmark.

\subsection{Lid Driven Cavity Problem}
\label{subsec:lid_driven_cavity}
As an extension of the convection problem, the second benchmark
problem given by Ghia \citep{ghia_high-re_1982} adds a driver for the
flow to introduce higher Reynolds numbers into the system, providing
more inertial force to overcome the viscous forces in the fluid. The
setup for this problem is equally simple, containing only the
Dirichlet conditions as given in Figure~\ref{fig:lid_driven_cavity}
and is only applied to the mass and momentum equations on the unit
square.
\begin{figure}[t!]
  \begin{center}
    \scalebox{1.5}{
      \input{chapters/nonlinear_problem/lid_driven_cavity.pdftex_t} }
  \end{center}
  \caption{\textbf{Problem setup for the lid driven cavity benchmark.}
    \textit{Dirichlet conditions of zero are set for the velocity on
      the left and right and bottom while the Dirichlet condition set
      on the top provides a driving force on the fluid.}}
  \label{fig:lid_driven_cavity}
\end{figure}
The top boundary condition will provide a driver for the flow and its
variation will in turn vary the Reynolds number of the fluid. An
increased velocity will generate more inertial forces in the fluid,
which will overcome the viscous forces and again increase the
influence of the nonlinear terms in Eq~(\ref{eq:ns_momentum}). Shadid
used Reynolds numbers up to \sn{1}{4} for this benchmark problem.

@@

@@

>>>>

\chapter{Monte Carlo Synthetic Acceleration Methods for the
  Navier-Stokes \\Equations\ }
\label{ch:nonlinear_problem}
Nonlinear transport problems are a common occurrence in single and
multiple physics problems in nuclear engineering. Systems of partial
differential equations such as those that describe fluid flow or more
general transport processes when discretized by conventional methods
yield discrete sets of stiff equations with nonlinearities present in
the variables. These sets of equations characterize momentum and
energy transport through diffusion, viscous forces, and inertial
forces. Traditionally, such systems have been solved by
linearizing them in a form where the nonlinearities in the variables
are eliminated and more traditional linear methods can be used for
solutions. Often characterized as segregated methods where physics
operators are split and their action on the system approximated in
steps, such methods lack consistency and accuracy in resolving the
nonlinear component of the solution. In the last 30 years, fully
consistent nonlinear methods based on Newton's method have become more
popular and many advances have been made in the nuclear engineering
field to employ these methods.

@@

\subsection{Newton-Krylov Methods\ }
\label{subsec:newton_krylov_methods}
A form of inexact Newton methods, \textit{Newton-Krylov methods} are
nonlinear iterative methods that leverage a Krylov subspace method as
the linear solver for generating the Newton correction
\citep{kelley_iterative_1995}. As outlined in
Appendix~\ref{ch:linear_problem}, Krylov methods are robust and enjoy
efficient parallel implementations on modern
architectures. Furthermore, their lack of explicit dependence on the
operator make them easier to implement than other
methods. Additionally, although many iterations can become memory
intensive due to the need to store the Krylov subspace for the
orthogonalization procedure, at each nonlinear iteration this cost is
reset as the Jacobian matrix will change due to its dependence on the
solution vector. This means that for every nonlinear iteration, a
completely new linear system is formed for generating the Newton
correction and we can modify the Krylov solver parameters accordingly
to accommodate this. In most nonlinear problems, the Jacobian operator
is generally non-symmetric and therefore either Krylov methods with
long recurrence relations that can handle non-symmetric systems must
be considered or the Newton correction system must be preconditioned
such that the operator is symmetric and short recurrence relation
methods can be potentially be used.

@@

\subsection{Thermal Convection Cavity Problem}
\label{subsec:natural_convection_cavity}
In 1983 a benchmark solution for the natural convection of air in a
square cavity was published \citep{de_vahl_davis_natural_1983} as
shown in Figure~\ref{fig:natural_convection_cavity} for the solution
of the energy, mass, and momentum equations.

\begin{figure}[t!]
  \begin{center}
    \scalebox{1.5}{
      \input{chapters/nonlinear_problem/natural_convection_cavity.pdftex_t} }
  \end{center}
  \caption{\textbf{Problem setup for the natural convection cavity
      benchmark.} \textit{Dirichlet conditions are set for the
      temperature on the left and right while Neumann conditions are
      set on the top and bottom of the Cartesian grid. The temperature
      gradients will cause buoyancy-driven flow. Zero velocity
      Dirichlet conditions are set on each boundary. No thermal source
      was present.}}
  \label{fig:natural_convection_cavity}
\end{figure}

In this problem, a rectilinear grid is applied to the unit square. No
heat flow is allowed out of the top and bottom of the square with a
zero Neumann condition specified. Buoyancy driven flow is generated by
the temperature gradient from the hot and cold Dirichlet conditions on
the left and right boundaries of the box. By adjusting the Rayleigh
number of the fluid (and therefore adjusting the ratio of convective
to conductive heat transfer), we can adjust the influence of the
nonlinear convection term in Eq~(\ref{eq:ns_momentum}). In Shadid's
work, Rayleigh numbers of up to \sn{1}{6} were used for this benchmark.

Temperature solutions for these equations are given in
Figure~\ref{fig:convection_isotherms} for Rayleigh numbers of
\sn{1}{3}, \sn{1}{4}, \sn{1}{5} and \sn{1}{6}. Note that as the
Rayleigh number is increased, the fluid begins to rotate in the
clockwise direction. This causes the temperature gradients to rotate
as well with the fluid, given the increased rotation observed for the
isotherms in the figure.

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/nonlinear_problem/convection_isotherms.png}
  \end{center}
  \caption{\textbf{Isotherms from solution to the thermal convection
        cavity problem.} \textit{Top left: Ra = \sn{1}{3}; Top right:
        Ra = \sn{1}{4}; Bottom left: Ra = \sn{1}{5}; Bottom left: Ra =
        \sn{1}{6}. Isotherms on unit temperature scale from 0 to 1
        with divisions of 0.1 for each calculation.}}
  \label{fig:convection_isotherms}
\end{figure}

\subsection{Lid Driven Cavity Problem}
\label{subsec:lid_driven_cavity}
As an extension of the convection problem, the second benchmark
problem given by Ghia \citep{ghia_high-re_1982} adds a driver for the
flow to introduce higher Reynolds numbers into the system, providing
more inertial force to overcome the viscous forces in the fluid. The
setup for this problem is equally simple, containing only the
Dirichlet conditions as given in Figure~\ref{fig:lid_driven_cavity}
and is only applied to the mass and momentum equations on the unit
square.

\begin{figure}[t!]
  \begin{center}
    \scalebox{1.5}{
      \input{chapters/nonlinear_problem/lid_driven_cavity.pdftex_t} }
  \end{center}
  \caption{\textbf{Problem setup for the lid driven cavity benchmark.}
    \textit{Dirichlet conditions of zero are set for the velocity on
      the left and right and bottom while the Dirichlet condition set
      on the top provides a driving force on the fluid.}}
  \label{fig:lid_driven_cavity}
\end{figure}

The top boundary condition will provide a driver for the flow and its
variation will in turn vary the Reynolds number of the fluid. An
increased velocity will generate more inertial forces in the fluid,
which will overcome the viscous forces and again increase the
influence of the nonlinear terms in Eq~(\ref{eq:ns_momentum}). Shadid
used Reynolds numbers up to \sn{1}{4} for this benchmark problem.

@@

For each of the benchmark problems, we will present the results of
computations using FANM and directly compare them to results using a
Newton-Krylov method in order to verify the correctness of FANM and
its applicability to fluid problems. For each case, we will compare
the global minimum, maximum, and average values for fluid velocity,
pressure, and temperature where applicable. In addition, we will
report the tolerance to which the nonlinear residual was converged.

Given the difficulty of these benchmark problems, the linear solve at
each Newton step was preconditioned with an algebraic multigrid
method. For FANM, this preconditioning was required to reduce the
spectral radius of the Jacobian operators generated at each Newton
step and for the Newton-Krylov solutions this preconditioning was
required to obtain good convergence properties although not necessary
for convergence. For every benchmark calculation, both FANM and
Newton-Krylov were preconditioned with the same multigrid parameters
such that the conditioning of the system would not be a factor in the
verification of the method or the perfomance comparison in the
following section.

\subsection{Thermal Convection Cavity Results}
\label{subsec:thermal_convection_verification}

\subsection{Lid Driven Cavity Results}
\label{subsec:lid_driven_verification}

@@

\subsection{Thermal Convection Cavity Results}
\label{subsec:thermal_convection_comparison}

\subsection{Lid Driven Cavity Results}
\label{subsec:lid_driven_comparison}
